---
title: 'Day 3: SoilR'
author: "Anna Tenberg"
date: "7/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SoilR and RothC model

### 1. Get familiar with the RothC model using the RothC guide (./Materials). 
#### How many pools are used in RothC and what is the assumption on splitting organic inputs into decomposable and resistant plant material? 

The four active compartments are Decomposable Plant Material (DPM), Resistant Plant Material (RPM), Microbial Biomass (BIO) and Humified Organic Matter (HUM).
The Assumption for splitting the organic input into DPM and RPM represents an influence of the landuse and vegetation type.

#### In how far is this assumption different from the assumption on the split between microbial biomass and humified substances? 

BIO and HUM split depends on the clay content of the soil, that gives input to the pore size and water capacity of the soil, which is an indicator for microbial activity. the influence is stronger and direct, instead of a generalization of many factors that are put into the ratio of DPM and RPM.

### 2. Run the RothC model using two different climate datasets (reference, scenario). As reference, repeat 1 year climate data of the Princeton dataset (see day 1) one hundred times. Use the climate scenario 2000-2100 you created on day 1. Describe the differences in soil C stocks in the C pools. Please note
  * a. Use the RothCModel function of the SoilR package
  * b. Assume a topsoil organic layer of 25 cm
  * c. Assume a soil organic carbon (SOC) of 69.7 Mg ha -1
  * d. Assume a clay content of 48%
  * e. Assume a C input of 2.7 Mg ha -1 yr -1
  * f. Assume a pan evaporation of Evp=data.frame(Month=1:12, Evp=c(12, 18, 35, 58, 82, 90, 97, 84, 54, 31, 14, 10)) for each year, regardless of scenario
  * g. Estimate inert organic matter (IOM parameter) based on Falloon method: FallIOM=0.049*SOC^(1.139)
  * h. Environmental conditions are considered by providing parameter xi. Calculate xi as the product of the temperature and moisture effect
  * i. Use the function fW.RothC for the moisture effect (here you need clay content and topsoil organic layer thickness provided above); use column b!
  * j. Use the function fT.RothC to calculate the temperature effect
  
```{r}
library(SoilR)
?RothCModel

#load reference and scenario from day1
reference <- read.delim("reference_2000-2100.txt")
rcp8.5 <- read.delim("rcp8.5_2000-2100.txt")

# provided parameter
Evp <- data.frame(Month=1:12, Evp=c(12, 18, 35, 58, 82, 90, 97, 84, 54, 31, 14, 10))
depth <- 25 #cm
SOC <- 69.7 #Mg ha -1
clay <- 48 #%
C_in <- 2.7 #Mg ha -1 yr -1
FallIOM <- 0.049*SOC^(1.139)

# calculation of the moisture decomposition rate
# for the reference data
P_ref <- reference$Precip[1:12]

fW_ref <- fW.RothC(P_ref, Evp$Evp, S.Thick = 23, pClay = 48, pE = 1, bare = FALSE)
fT_ref <- fT.RothC(Temp = reference$Temp[1:12])

xi_ref <- fW_ref$b * fT_ref # df mit zeit und produkt
xi_ref <- data.frame("month" = 1:12, xi)

# multiply xi_ref to 100 years
xi_100 <- xi_ref
for (i in 1:100) {
  xi_100 <- rbind(xi_100, xi_ref)
}

xi_100$month <- 1:length(xi_100$month)

# for the simulation data
P_rcp <- rcp8.5$Precip

Evp100 <- Evp$Evp
for (i in 1:100) {
  Evp100 <- cbind(Evp100, Evp$Evp)
}

fW_rcp <- fW.RothC(P_rcp, Evp100, S.Thick = 23, pClay = 48, pE = 1, bare = FALSE)
fT_rcp <- fT.RothC(Temp = rcp8.5$Temp)

xi_rcp <- fW_rcp$b * fT_rcp
xi_rcp <- cbind(1:length(xi_rcp), xi_rcp)

# run of the model
ref_model <- RothCModel(1:length(xi_100$month), ks = c(k.DPM = 10, k.RPM = 0.3, k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), In = C_in, DR = 1.44, clay = 48, xi = xi_100, solver = deSolve.lsoda.wrapper, pass = FALSE)

# carbonstock of reference data
cstock_ref <- getC(ref_model)

rcp_model <- RothCModel(1:length(xi_100$month), ks = c(k.DPM = 10, k.RPM = 0.3, k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), In = C_in, DR = 1.44, clay = 48, xi = as.data.frame(xi_rcp), solver = deSolve.lsoda.wrapper, pass = FALSE)

# carbonstock of model
cstock_rcp <- getC(rcp_model)

```

### 3. Decomposition rates are a fundamental parameter for soil carbon models. They can be obtained from laboratory experiments. Use the dataset “./Materials/00_experimentalData.txt” of cumulative CO2 over time to fit a one-pool model and a two-pool model. Create a function that runs the SoilR function “TwopFeedbackModel” and “OnepModel” with one parameters vector. The function has to return the CO2 released from all pools involved (read on “getAccumulatedRelease”). For the optimization of the parameters, use the modFit function of the FME package. Be aware that modFit takes a cost function that is optimized. It is probably the most convenient to define a costfunction (costf) that has one argument (parameter vector) and calculates the model output based on the parameter vector and returns the values returned from the FME function modCost. Plot the cumulative CO2 release for both models. Why does it make sense to have multi-pool models?

```{r, eval= FALSE}
exp_data <- read.delim("00_experimentalData.txt")

### doesnt run, just a sketch overview, due to not enough time 
pool_model <- function(parameter, onepool = TRUE){
  if(onepool == TRUE){
    model <- OnepModel(t, k, C0, In, xi = 1, solver = deSolve.lsoda.wrapper, pass = FALSE)
  } else {
    model <- TwopFeedbackModel(t = , ks = c(pool1, pool2), a21, a12, C0 = c(0, 0, 0, 0, FallIOM), In = C_in, xi = xi_run, solver = deSolve.lsoda.wrapper, pass = FALSE)
  }
  return(getAccumulatedRelease(model))
}

costf <- function(parameter, onepool = TRUE){
  model <- pool_model(parameter, onepool = onepool)
  mincost <- modCost(model, obs)
  return(mincost)
}

library(FME)
opt_CO2 <- modFit(costf, p, ..., lower = exp_data$cummCO2[1], upper = exp_data$cummCO2[21], method = c("Marq", "Port", "Newton",            "Nelder-Mead", "BFGS", "CG", "L-BFGS-B", "SANN", "Pseudo", "bobyqa"), jac = NULL, control = list(), hessian = TRUE)

t <- exp_data$time
k <- exp_data$cummCO2

```

