---
title: 'Day 3: SoilR'
author: "Anna Tenberg"
date: "7/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SoilR and RothC model

### 1. Get familiar with the RothC model using the RothC guide (./Materials). 
#### How many pools are used in RothC and what is the assumption on splitting organic inputs into decomposable and resistant plant material? 

The four active compartments are Decomposable Plant Material (DPM), Resistant Plant Material (RPM), Microbial Biomass (BIO) and Humified Organic Matter (HUM).
The Assumption for splitting the organic input into DPM and RPM represents an influence of the landuse and vegetation type.

#### In how far is this assumption different from the assumption on the split between microbial biomass and humified substances? 

BIO and HUM split depends on the clay content of the soil, that gives input to the pore size and water capacity of the soil, which is an indicator for microbial activity. the influence is stronger and direct, instead of a generalization of many factors that are put into the ratio of DPM and RPM.

### 2. Run the RothC model using two different climate datasets (reference, scenario). As reference, repeat 1 year climate data of the Princeton dataset (see day 1) one hundred times. Use the climate scenario 2000-2100 you created on day 1. Describe the differences in soil C stocks in the C pools. Please note
  * a. Use the RothCModel function of the SoilR package
  * b. Assume a topsoil organic layer of 25 cm
  * c. Assume a soil organic carbon (SOC) of 69.7 Mg ha -1
  * d. Assume a clay content of 48%
  * e. Assume a C input of 2.7 Mg ha -1 yr -1
  * f. Assume a pan evaporation of Evp=data.frame(Month=1:12, Evp=c(12, 18, 35, 58, 82, 90, 97, 84, 54, 31, 14, 10)) for each year, regardless of scenario
  * g. Estimate inert organic matter (IOM parameter) based on Falloon method: FallIOM=0.049*SOC^(1.139)
  * h. Environmental conditions are considered by providing parameter xi. Calculate xi as the product of the temperature and moisture effect
  * i. Use the function fW.RothC for the moisture effect (here you need clay content and topsoil organic layer thickness provided above); use column b!
  * j. Use the function fT.RothC to calculate the temperature effect
  
  
```{r}
refdata <- read.table("materials/reference_2000-2100.txt", header=T)
scenariodata <- read.table("materials/rcp8.5_2000-2100.txt", header=T)
```


Add evaporation:

```{r}
Evp=data.frame(Month=1:12, Evp=c(12, 18, 35, 58, 82, 90, 97, 84, 54, 31, 14, 10))
Evp2 <- Evp

for (i in 1:100) {
  Evp <- rbind(Evp, Evp2)
}

refdata <- cbind(refdata, Evp$Evp)
scenariodata <- cbind(scenariodata, Evp$Evp)

```


Initial IOM:

```{r}
FallIOM=0.049* 69.7^(1.139) # 69.7 is the SOM
```


Calculation of the moisture effect with the function fW.RothC

```{r}
#?fW.RothC
# pE is evaporation coefficient and here set to 1

# for the reference data
moisture_ref <- fW.RothC(refdata$Precip, refdata$`Evp$Evp`, S.Thick = 25, 
                         pClay = 48, pE = 1, bare = FALSE) 

# for the scenario data
moisture_sc <- fW.RothC(scenariodata$Precip, scenariodata$`Evp$Evp`, S.Thick = 25, 
                        pClay = 48, pE = 1, bare = FALSE) 


```


Calculation of the temperature effect with the function fT.RothC

```{r}
#?fT.RothC

# reference temp effect
tempeff_ref <- fT.RothC(refdata$Temp)

# scenario temp effect
tempeff_sc <- fT.RothC(scenariodata$Temp)


```


Xi is the product of the moisture effect and the temperature effect

```{r}
# reference
xi_ref = data.frame("month"=1:1212 , "xi"=moisture_ref$b*tempeff_ref)

#scenario
xi_sc = data.frame("month"=1:1212 , "xi"=moisture_sc$b*tempeff_sc)


#?RothCModel
# DPM, RPM, BIO and HUM initial 0
```

Model for the reference data:

```{r}

refmodel <- RothCModel(xi_ref$month, ks = c(k.DPM = 10, k.RPM = 0.3, 
            k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), 
            In = 2.7, DR = 1.44, clay = 48, xi = xi_ref, solver = deSolve.lsoda.wrapper, pass = FALSE)

cstockref_ref <- getC(refmodel) # get the pools
#head(cstockref_ref) # in the beginning
#tail(cstockref_ref) # in the end

```


```{r}

# Reference data: Stock changes
plot(cstockref_ref[,1]) # DPM
plot(cstockref_ref[,2]) # RPM
plot(cstockref_ref[,3]) # BIO
plot(cstockref_ref[,4]) # HUM, needs ~ 360 month = 30 years
plot(cstockref_ref[,5]) # IoM # does not change at all


```

Model for the scenario data:

```{r}

# scenario data

scmodel <- RothCModel(xi_sc$month, ks = c(k.DPM = 10, k.RPM = 0.3, 
           k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), 
           In = 2.7, DR = 1.44, clay = 48, xi = xi_sc, solver = deSolve.lsoda.wrapper, pass = FALSE)

cstockref_sc <- getC(scmodel)
#head(cstockref_sc)
#tail(cstockref_sc)
```


```{r}
# Scenario data: Stock changes
plot(cstockref_sc[,1]) 
plot(cstockref_sc[,2])
plot(cstockref_sc[,3])
plot(cstockref_sc[,4])
#plot(cstockref_sc[,5])

```

## Spin up
+ 30 years of reference data (360 months)

```{r}
# Reference data
spinned_up_refdata <- rbind(refdata[1:360,], refdata)

# Scenariodata
spinned_up_scenariodata <- rbind(refdata[1:360,], scenariodata)

```

Moisture effect for spinned up data

```{r}

# for the reference data
spinned_moisture_ref <- fW.RothC(spinned_up_refdata$Precip, spinned_up_refdata$`Evp$Evp`, 
                                 S.Thick = 25, pClay = 48, pE = 1, bare = FALSE) # pE is pan evaporation

# for the scenario data
spinned_moisture_sc <- fW.RothC(spinned_up_scenariodata$Precip, spinned_up_scenariodata$`Evp$Evp`, 
                                S.Thick = 25, pClay = 48, pE = 1, bare = FALSE) # pE is pan evaporation


```


Temperature effect for spinned up data

```{r}
#?fT.RothC

# reference temp effect
spinned_tempeff_ref <- fT.RothC(spinned_up_refdata$Temp)

# scenario temp effect
spinned_tempeff_sc <- fT.RothC(spinned_up_scenariodata$Temp)


```

xi for spinned up data

```{r}
xi_ref_spinned = data.frame("month"=1:1572 , "xi"=spinned_moisture_ref$b*spinned_tempeff_ref)

xi_sc_spinned = data.frame("month"=1:1572 , "xi"=spinned_moisture_sc$b*spinned_tempeff_sc)


#?RothCModel

# spinned up reference model

spinned_refmodel <- RothCModel(xi_ref_spinned$month, ks = c(k.DPM = 10, k.RPM = 0.3, 
                    k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), 
                    In = 2.7, DR = 1.44, clay = 48, xi = xi_ref_spinned, 
                    solver = deSolve.lsoda.wrapper, pass = FALSE)

spinned_cstockref_ref <- getC(spinned_refmodel)
#head(spinned_cstockref_ref)
#tail(spinned_cstockref_ref)


# spinned up scenario data

spinned_scmodel <- RothCModel(xi_sc_spinned$month, ks = c(k.DPM = 10, k.RPM = 0.3, 
                   k.BIO = 0.66, k.HUM = 0.02, k.IOM = 0), C0 = c(0, 0, 0, 0, FallIOM), 
                   In = 2.7, DR = 1.44, clay = 48, xi = xi_sc_spinned, 
                   solver = deSolve.lsoda.wrapper, pass = FALSE)

spinned_cstockref_sc <- getC(scmodel)
#head(spinned_cstockref_sc)
#tail(spinned_cstockref_sc)

```

Now, there should be no changes in time for the different stocks in the reference data if we begin at the end of the spin up.


```{r}

plot(spinned_cstockref_ref[360:nrow(spinned_cstockref_ref),1]) # DPM
plot(spinned_cstockref_ref[360:nrow(spinned_cstockref_ref),2]) # RPM
plot(spinned_cstockref_ref[360:nrow(spinned_cstockref_ref),3]) # BIO
plot(spinned_cstockref_ref[360:nrow(spinned_cstockref_ref),4]) # HUM, only very small changes seen now
#plot(spinned_cstockref_ref[360:nrow(spinned_cstockref_ref),5]) # IoM


```


Let's look at the scenario data.

```{r}

plot(spinned_cstockref_sc[360:nrow(spinned_cstockref_sc),1]) # DPM
plot(spinned_cstockref_sc[360:nrow(spinned_cstockref_sc),2]) # RPM
plot(spinned_cstockref_sc[360:nrow(spinned_cstockref_sc),3]) # BIO
plot(spinned_cstockref_sc[360:nrow(spinned_cstockref_sc),4]) # HUM
#plot(spinned_cstockref_sc[360:nrow(spinned_cstockref_sc),5]) # IoM
```

Changes in climate change scenario also after the spin up!


### 3. Decomposition rates are a fundamental parameter for soil carbon models. They can be obtained from laboratory experiments. Use the dataset “./Materials/00_experimentalData.txt” of cumulative CO2 over time to fit a one-pool model and a two-pool model. Create a function that runs the SoilR function “TwopFeedbackModel” and “OnepModel” with one parameters vector. The function has to return the CO2 released from all pools involved (read on “getAccumulatedRelease”). For the optimization of the parameters, use the modFit function of the FME package. Be aware that modFit takes a cost function that is optimized. It is probably the most convenient to define a costfunction (costf) that has one argument (parameter vector) and calculates the model output based on the parameter vector and returns the values returned from the FME function modCost. Plot the cumulative CO2 release for both models. Why does it make sense to have multi-pool models?
